---
title: "Event_Studies"
author: "Jiaxi Li"
date: "June 2, 2019"
output: 
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
# remove the error messages and warnings in the project
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Load the useful packages
library(tidyverse)
library(here)
library(dplyr)
theme_set(theme_minimal())
set.seed(1234)
```

I would first obtain the PERMNO's and returns and share outstandings from CRSP Monthly Stock from 1959 to 2018. The conditional statements are share code 10 or 11, exchange code from 1 to 3, number of shares outstanding is not null and holding period return is not null.

```{r, eval = FALSE}
# The dataset is crsp_monthly.csv
Stocks_monthly <- read_csv(here("data","crsp_monthly.csv")) %>%
  mutate(yearmonth = floor(date/100))

# Save and write PERMNOs
# PERMNOs <- unique(Stocks_montly$PERMNO)
# write.table(PERMNOs, file = "Permno.txt", sep = "\t", row.names = FALSE, col.names = FALSE)
```

After extracting all the PERMNOs, I extract the book values for the stocks with the book value from the PERMNOs. The conditional Statement is Book value per Share is not null.

```{r, eval = FALSE}
# The dataset is book_value.csv
Book <- read_csv(here("data","book_value.csv")) %>%
  mutate(yearmonth = floor(datadate/100)) %>%
  rename(PERMNO = LPERMNO)

# Save and write PERMNOs2
# PERMNOs2 <- unique(Book$LPERMNO)
# PERMNOs2 = PERMNOs2[order(PERMNOs2)]
# write.table(PERMNOs2, file = "Permno2.txt", sep = "\t", row.names = FALSE, col.names = FALSE)
```

Caculate the size, book-to-market, and momemtum for each stocks. The book equity value is the book equity value from the most recent fiscal year. Here, I assume that each of the value is constant for a given month.

```{r, eval = FALSE}
# merge stocks_monthly and book by PERMNO and yearmonth
portfolios <- left_join(Stocks_monthly, Book, by = c('PERMNO','yearmonth')) %>%
  select(PERMNO, PRC, RET, SHROUT, yearmonth, fyear, fyr, bkvlps) %>%
# convert to log return, calculate the size and book to market and momentum (return from 12 months before to 2 months before)
  mutate(PRC = abs(PRC),
         size = PRC * SHROUT,
         RET = log(RET+1))

yearmonths <- unique(portfolios$yearmonth)
yearmonths <- data.frame( yearmonth = yearmonths[order(yearmonths)]) %>%
  mutate(lag12_ym = lag(yearmonth, 12))

# create the momentum and other factors
portfolios2 <- portfolios %>%
  group_by(PERMNO) %>%
  mutate(MOM = reduce(map(2:12, ~ lag(RET, ., 0)), `+`)，
         lag12 = lag(yearmonth,12)) %>%
  left_join(yearmonths, by = 'yearmonth') %>%
  mutate(fyear_1 = lag(fyear,1))

# fill in the missing fiscal year
for (i in 1:14) {
  temp <- lag(portfolios2$fyear_1,1)
  portfolios2$fyear_1[is.na(portfolios2$fyear_1)] <- temp[is.na(portfolios2$fyear_1)]
}

# extract book value
book_value <- portfolios2 %>%
  select(PERMNO, fyear, bkvlps) %>%
  filter(!is.na(bkvlps)) %>%
  rename(fyear_1 = fyear,
         book = bkvlps)
  
# merge in the book values
portfolios2 <- portfolios2 %>%
  left_join(book_value, by = c('fyear_1','PERMNO'))

# Calculate Book to Market (BTM)
portfolios2 <- portfolios2 %>%
  mutate(BTM = PRC/book) %>%
  filter(!is.na(lag12),
         lag12 == lag12_ym) %>%
  select(PERMNO, yearmonth, size, BTM, MOM) %>% na.omit()

# form the quintiles for size BTM and MOM
quintiles <- portfolios2 %>%
  group_by(yearmonth) %>%
  summarize(size_1 = quantile(size, 0.2),
            size_2 = quantile(size, 0.4),
            size_3 = quantile(size, 0.6),
            size_4 = quantile(size, 0.8),
            BTM_1 = quantile(BTM, 0.2),
            BTM_2 = quantile(BTM, 0.4),
            BTM_3 = quantile(BTM, 0.6),
            BTM_4 = quantile(BTM, 0.8),
            MOM_1 = quantile(MOM, 0.2),
            MOM_2 = quantile(MOM, 0.4),
            MOM_3 = quantile(MOM, 0.6),
            MOM_4 = quantile(MOM, 0.8))

# merge the quintiles data with the portfolios2
portfolios3 <- left_join(portfolios2, quintiles) %>%
  mutate(size_group = 1 + 1*(size>size_1) + 1*(size>size_2) + 1*(size>size_3) + 1*(size>size_4),
         BTM_group = 1 + 1*(BTM>BTM_1) + 1*(BTM>BTM_2) + 1*(BTM>BTM_3) + 1*(BTM>BTM_4),
         MOM_group = 1 + 1*(MOM>MOM_1) + 1*(MOM>MOM_2) + 1*(MOM>MOM_3) + 1*(MOM>MOM_4)) %>%
  select(PERMNO, yearmonth, size_group, BTM_group, MOM_group) %>%
  rename(yearmonth_1 = yearmonth)

# create the matching yearmonths
yearmonths <- yearmonths %>%
  select(yearmonth) %>%
  mutate(yearmonth_1 = lag(yearmonth,1))

# merge the months
portfolios3 <- portfolios3 %>%
  left_join(yearmonths) %>%
  select(-yearmonth_1)
```

After defining the groups for each PERMNO/month, One needs to remember that when one uses lagged characteristics to find the matching portfolios.

Extract the Permnos again and use them to obtain the dividend ex date (not null; from CRSP/Compustat Merged Database - Security Daily) and earnings announcement date (not null; CRSP/Compustat Merged Database - Fundamentals Quarterly) and the daily returns (not null; CRSP Daily Stock)

```{r, eval = FALSE}
# load daily returns
daily <- read_csv(here("data","crsp_daily.csv")) %>%
  select(-c("SHRCD","EXCHCD")) %>%
  mutate(yearmonth = floor(date/100)) %>%
  inner_join(portfolios3)
  
# calculate the average returns for each porfolio.
ports <- daily %>%
  group_by(date, size_group, BTM_group, MOM_group) %>%
  summarize(mean_ret = mean(RET))

# merge all the daily returns with the portfolio return
daily <- daily %>%
  inner_join(ports) %>%
  select(PERMNO, date, RET, mean_ret)

# load dividend
dividend <- read_csv(here("data","dividend.csv")) %>%
  rename(PERMNO = LPERMNO,
         date = datadate) %>%
  select(PERMNO, date) %>%
  mutate(div_event = 1)

# load earnings announcement
earnings <- read_csv(here("data","earnings.csv")) %>%
  rename(PERMNO = LPERMNO,
         date = rdq) %>%
  select(PERMNO, date) %>%
  mutate(earn_event = 1)

# merge all return, dividend and earnings annoucenment data and compute abnormal returns
daily <- daily %>%
  left_join(dividend) %>%
  left_join(earnings) %>%
  mutate(AR = RET - mean_ret) %>%
  select(-c(RET, mean_ret))

write_csv(daily, "PS2.csv")
```

Save all the daily Abnormal Returns and dividend and earnings announcement events into a data file PS2.csv
```{r}
daily <- read_csv(here("PS2.csv"))

# Extract the -10 to 10 day window for the dividend ex date
divAR <- data.frame(date = -10:10, AR = -10:10, CAR = -10:10)
for (i in -10:10) {
  divAR$AR[11+i] <- daily %>%
    group_by(PERMNO) %>%
    select(!is.na(lag()))
}
divAR <- daily %>%
  group_by(PERMNO)
```


# Session info

```{r}
session_info()
```
