---
title: "Event_Studies"
author: "Jiaxi Li"
date: "June 2, 2019"
output: 
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
# remove the error messages and warnings in the project
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Load the useful packages
library(tidyverse)
library(here)
library(dplyr)
theme_set(theme_minimal())
set.seed(1234)
```

I would first obtain the PERMNO's and returns and share outstandings from CRSP Monthly Stock from 1959 to 2018. The conditional statements are share code 10 or 11, exchange code from 1 to 3, number of shares outstanding is not null and holding period return is not null.

```{r}
# The dataset is crsp_monthly.csv
Stocks_montly <- read_csv(here("data","crsp_monthly.csv")) %>%
  mutate(yearmonth = floor(date/100))

# Save and write PERMNOs
# PERMNOs <- unique(Stocks_montly$PERMNO)
# write.table(PERMNOs, file = "Permno.txt", sep = "\t", row.names = FALSE, col.names = FALSE)
```

After extracting all the PERMNOs, I extract the book values for the stocks with the book value from the PERMNOs. The conditional Statement is Book value per Share is not null.

```{r}
# The dataset is book_value.csv
Book <- read_csv(here("data","book_value.csv")) %>%
  mutate(yearmonth = floor(datadate/100))

# Save and write PERMNOs2
# PERMNOs2 <- unique(Book$LPERMNO)
# PERMNOs2 = PERMNOs2[order(PERMNOs2)]
# write.table(PERMNOs2, file = "Permno2.txt", sep = "\t", row.names = FALSE, col.names = FALSE)
```

Extract the Permnos again and use them to obtain the dividend ex date (not null; from CRSP/Compustat Merged Database - Security Daily) and earnings announcement date (not null; CRSP/Compustat Merged Database - Fundamentals Quarterly) and the daily returns (not null; CRSP Daily Stock)


# Session info

```{r}
session_info()
```
