---
title: "Event_Studies"
author: "Jiaxi Li"
date: "June 2, 2019"
output: 
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
# remove the error messages and warnings in the project
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# Load the useful packages
library(tidyverse)
library(here)
library(dplyr)
theme_set(theme_minimal())
set.seed(1234)
```

# Data Cleaning and Variable Calculation
I would first obtain the PERMNO's and returns and share outstandings from CRSP Monthly Stock from 1959 to 2018. The conditional statements are share code 10 or 11, exchange code from 1 to 3, number of shares outstanding is not null and holding period return is not null.

```{r, eval = FALSE}
# The dataset is crsp_monthly.csv
Stocks_monthly <- read_csv(here("data","crsp_monthly.csv")) %>%
  mutate(yearmonth = floor(date/100))

# Save and write PERMNOs
# PERMNOs <- unique(Stocks_montly$PERMNO)
# write.table(PERMNOs, file = "Permno.txt", sep = "\t", row.names = FALSE, col.names = FALSE)
```

After extracting all the PERMNOs, I extract the book values for the stocks with the book value from the PERMNOs. The conditional Statement is Book value per Share is not null.

```{r, eval = FALSE}
# The dataset is book_value.csv
Book <- read_csv(here("data","book_value.csv")) %>%
  mutate(yearmonth = floor(datadate/100)) %>%
  rename(PERMNO = LPERMNO)

# Save and write PERMNOs2
# PERMNOs2 <- unique(Book$LPERMNO)
# PERMNOs2 = PERMNOs2[order(PERMNOs2)]
# write.table(PERMNOs2, file = "Permno2.txt", sep = "\t", row.names = FALSE, col.names = FALSE)
```

Caculate the size, book-to-market, and momemtum for each stocks. The book equity value is the book equity value from the most recent fiscal year. Here, I assume that each of the value is constant for a given month.

```{r, eval = FALSE}
# merge stocks_monthly and book by PERMNO and yearmonth
portfolios <- left_join(Stocks_monthly, Book, by = c('PERMNO','yearmonth')) %>%
  select(PERMNO, PRC, RET, SHROUT, yearmonth, fyear, fyr, bkvlps) %>%
# convert to log return, calculate the size and book to market and momentum (return from 12 months before to 2 months before)
  mutate(PRC = abs(PRC),
         size = PRC * SHROUT,
         RET = log(RET+1))

yearmonths <- unique(portfolios$yearmonth)
yearmonths <- data.frame( yearmonth = yearmonths[order(yearmonths)]) %>%
  mutate(lag12_ym = lag(yearmonth, 12))

# create the momentum and other factors
portfolios2 <- portfolios %>%
  group_by(PERMNO) %>%
  mutate(MOM = reduce(map(2:12, ~ lag(RET, ., 0)), `+`)，
         lag12 = lag(yearmonth,12)) %>%
  left_join(yearmonths, by = 'yearmonth') %>%
  mutate(fyear_1 = lag(fyear,1))

# fill in the missing fiscal year
for (i in 1:14) {
  temp <- lag(portfolios2$fyear_1,1)
  portfolios2$fyear_1[is.na(portfolios2$fyear_1)] <- temp[is.na(portfolios2$fyear_1)]
}

# extract book value
book_value <- portfolios2 %>%
  select(PERMNO, fyear, bkvlps) %>%
  filter(!is.na(bkvlps)) %>%
  rename(fyear_1 = fyear,
         book = bkvlps)
  
# merge in the book values
portfolios2 <- portfolios2 %>%
  left_join(book_value, by = c('fyear_1','PERMNO'))

# Calculate Book to Market (BTM)
portfolios2 <- portfolios2 %>%
  mutate(BTM = PRC/book) %>%
  filter(!is.na(lag12),
         lag12 == lag12_ym) %>%
  select(PERMNO, yearmonth, size, BTM, MOM) %>% na.omit()

# form the quintiles for size BTM and MOM
quintiles <- portfolios2 %>%
  group_by(yearmonth) %>%
  summarize(size_1 = quantile(size, 0.2),
            size_2 = quantile(size, 0.4),
            size_3 = quantile(size, 0.6),
            size_4 = quantile(size, 0.8),
            BTM_1 = quantile(BTM, 0.2),
            BTM_2 = quantile(BTM, 0.4),
            BTM_3 = quantile(BTM, 0.6),
            BTM_4 = quantile(BTM, 0.8),
            MOM_1 = quantile(MOM, 0.2),
            MOM_2 = quantile(MOM, 0.4),
            MOM_3 = quantile(MOM, 0.6),
            MOM_4 = quantile(MOM, 0.8))

# merge the quintiles data with the portfolios2
portfolios3 <- left_join(portfolios2, quintiles) %>%
  mutate(size_group = 1 + 1*(size>size_1) + 1*(size>size_2) + 1*(size>size_3) + 1*(size>size_4),
         BTM_group = 1 + 1*(BTM>BTM_1) + 1*(BTM>BTM_2) + 1*(BTM>BTM_3) + 1*(BTM>BTM_4),
         MOM_group = 1 + 1*(MOM>MOM_1) + 1*(MOM>MOM_2) + 1*(MOM>MOM_3) + 1*(MOM>MOM_4)) %>%
  select(PERMNO, yearmonth, size_group, BTM_group, MOM_group) %>%
  rename(yearmonth_1 = yearmonth)

# create the matching yearmonths
yearmonths <- yearmonths %>%
  select(yearmonth) %>%
  mutate(yearmonth_1 = lag(yearmonth,1))

# merge the months
portfolios3 <- portfolios3 %>%
  left_join(yearmonths) %>%
  select(-yearmonth_1)
```

After defining the groups for each PERMNO/month, One needs to remember that when one uses lagged characteristics to find the matching portfolios.

Extract the Permnos again and use them to obtain the dividend ex date (not null; from CRSP/Compustat Merged Database - Security Daily) and earnings announcement date (not null; CRSP/Compustat Merged Database - Fundamentals Quarterly) and the daily returns (not null; CRSP Daily Stock)

```{r, eval = FALSE}
# load daily returns
daily <- read_csv(here("data","crsp_daily.csv")) %>%
  select(-c("SHRCD","EXCHCD")) %>%
  mutate(yearmonth = floor(date/100)) %>%
  inner_join(portfolios3)
  
# calculate the average returns for each porfolio.
ports <- daily %>%
  group_by(date, size_group, BTM_group, MOM_group) %>%
  summarize(mean_ret = mean(RET))

# merge all the daily returns with the portfolio return
daily <- daily %>%
  inner_join(ports) %>%
  select(PERMNO, date, RET, mean_ret)

# load dividend
dividend <- read_csv(here("data","dividend.csv")) %>%
  rename(PERMNO = LPERMNO,
         date = datadate) %>%
  select(PERMNO, date) %>%
  mutate(div_event = 1)

# load earnings announcement
earnings <- read_csv(here("data","earnings.csv")) %>%
  rename(PERMNO = LPERMNO,
         date = rdq) %>%
  select(PERMNO, date) %>%
  mutate(earn_event = 1)

# merge all return, dividend and earnings annoucenment data and compute abnormal returns
daily <- daily %>%
  left_join(dividend) %>%
  left_join(earnings) %>%
  mutate(AR = RET - mean_ret) %>%
  select(-c(RET, mean_ret)) %>%
  ungroup()

write_csv(daily, "PS2.csv")
```

Save all the daily Abnormal Returns and dividend and earnings announcement events into a data file PS2.csv

# The Dividend Event
```{r}
daily <- read_csv(here("PS2.csv"))

# Extract the -10 to 10 day window for the dividend ex date
divAR <- data.frame(date = -10:10, AR = -10:10, CAR = -10:10)
for (i in -10:-1) {
  divAR$AR[11+i] <- daily %>%
    group_by(PERMNO) %>%
    mutate(lag_event = lead(div_event,-i)) %>%
    filter(!is.na(lag_event)) %>%
    ungroup() %>%
    summarise(AAR = mean(AR)) %>%
    as.numeric()
}
divAR$AR[11] <- daily %>%
  group_by(PERMNO) %>%
  filter(!is.na(div_event)) %>%
  ungroup() %>%
  summarise(AAR = mean(AR)) %>%
  as.numeric()
for (i in 1:10) {
  divAR$AR[11+i] <- daily %>%
    group_by(PERMNO) %>%
    mutate(lag_event = lag(div_event,i)) %>%
    filter(!is.na(lag_event)) %>%
    ungroup() %>%
    summarise(AAR = mean(AR)) %>%
    as.numeric()
}

divAR <- divAR %>%
  mutate(CAR = cumsum(AR))

# plot the AR and CAR
divAR %>%
  ggplot() +
  geom_bar(aes(x = date, y = AR*100), stat = "identity") +
  geom_line(aes(x = date, y = CAR*100/2)) +
  geom_point(aes(x = date, y = CAR*100/2), fill = "white") +
  scale_y_continuous(
    name = expression("AR (%)"), 
    sec.axis = sec_axis(~ .*2 ,name = "CAR (%)")) +
  labs(title = "AR and CAR around Dividend ex Date",
       caption = "Data Source: The Center for Research in Security Prices",
       x = "Event date") +
  theme(plot.title = element_text(hjust=0.5))
```

It seems that there is a positive abnormal return for dividend event. The graph supports the findings of Hartzmark and Solomon (2013): "The premium is consistent with price pressure from dividend-seeking investors. Measures of liquidity and demand for dividends are associated with larger price increases in the period before the ex-day (when there is no news about the dividend), and larger reversals afterwards."


# The Earnings Announcement Event
```{r}
# Extract the -10 to 10 day window for earnings announcement date
earnAR <- data.frame(date = -10:10, AR = -10:10, CAR = -10:10)
for (i in -10:-1) {
  earnAR$AR[11+i] <- daily %>%
    group_by(PERMNO) %>%
    mutate(lag_event = lead(earn_event,-i)) %>%
    filter(!is.na(lag_event)) %>%
    ungroup() %>%
    summarise(AAR = mean(AR)) %>%
    as.numeric()
}
earnAR$AR[11] <- daily %>%
  group_by(PERMNO) %>%
  filter(!is.na(earn_event)) %>%
  ungroup() %>%
  summarise(AAR = mean(AR)) %>%
  as.numeric()
for (i in 1:10) {
  earnAR$AR[11+i] <- daily %>%
    group_by(PERMNO) %>%
    mutate(lag_event = lag(earn_event,i)) %>%
    filter(!is.na(lag_event)) %>%
    ungroup() %>%
    summarise(AAR = mean(AR)) %>%
    as.numeric()
}

earnAR <- earnAR %>%
  mutate(CAR = cumsum(AR))

# plot the AR and CAR
earnAR %>%
  ggplot() +
  geom_bar(aes(x = date, y = AR*100), stat = "identity") +
  geom_point(aes(x = date, y = CAR*100/2), fill = "white") +
  geom_line(aes(x = date, y = CAR*100/2)) +
  scale_y_continuous(
    name = expression("AR (%)"), 
    sec.axis = sec_axis(~ .*2 ,name = "CAR (%)")) +
  labs(title = "AR and CAR around Earnings Announcement",
       caption = "Data Source: The Center for Research in Security Prices",
       x = "Event date") +
  theme(plot.title = element_text(hjust=0.5))

```

It seems that there is a positive abnormal return for earnings annoucement as well. The effect of positive Abnormal return is gradually increasing and there is no jump as dividend ex date. It seems market are updating information before the event. This confirms the claim by Frazzini and Lamont (2007): "On average, stock prices rise around scheduled earnings announcement dates. We show that this earnings announcement premium is large, robust, and strongly related to the fact that volume surges around announcement dates.""

# Session info

```{r}
devtools::session_info()
```
